Discovering causal factors with back-to-back regression

### 2019-05-22 paper/meg_exps.tex
wip: meg

\subsection{The MOUS dataset}

Next, we apply our method to brain imaging data, on the ``Mother Of all Unification
Studies'' (MOUS) dataset \cite{schoffelen2019204}. The dataset consists of
anonymized multimodal neuroimaging data that has been acquired from 204 healthy
human subjects, which participated in either a auditory or a visual version of a
language experiment. In this work, we use the subset corresponding to the visual
variant recorded using magneto-encephalography (MEG).

Subjects were presented a set of 120 sentences in Dutch, and scrambled lists
of the same words for a controlled experiment. Each word was presented on a
screen for no less than 300 ms and no more than 1400ms, 351 ms on average.
Successive words were separated by an empty screen for 300ms, and successive
sentences were separated by an empty screen for a few (3-4) seconds.

The data contains the timestamp for each word presentation, allowing to align
time series to stimuli onsets for all words.

As a result, we obtain, for each of the 301 channels (273 magnetometers on the
subject and 28 compensation channels measuring the room) and each word
presented, a time series of 67 real-values spanning $\left[-100ms, 1000ms\right]$ relative to
the stimulus onset. Each of the 102 subjects (we used data from 93 subjects) was
presented approximately 2700 words during the experiment.

We show, in Figure \ref{fig:megavg}, the average response across words of the 
magnetometers for the first subject. Time $t=0$ corresponds to the stimulus onset.

@@

\subsection{Application of the method}

Using this dataset, we study which features of input words can elicit a strong
MEG response in the brain.  We add to each word (i) visual features: word length
(1 dim.), letter count (26 dim.), and (ii) lexical features : word frequency (1
dim.) part-of-speech tag (one of ADJ, ADP, ADV, CONJ, DET, NOUN, NUM, PRON,
PROPN, VERB; 10 dim.). Sentence-level word features are left for future study.

This procedure yields an $\left(n \times 38\right)$-dimensional input matrix $X$,
where $n$ is the number of words for the subject. Each feature in $X$ is centered and
scaled for zero mean and unit variance.

>>>>

\subsection{Dataset}

Next, we apply our method to brain imaging data from the anonymized multimodal
neuroimaging ``Mother Of all Unification Studies'' (MOUS) dataset \cite{schoffelen2019204}.
The dataset contains magneto-encephalography (MEG) recordings of 102 healthy native-Dutch
adults who participated in a reading task.

Subjects were exposed to a rapid serial visual presentation of Dutch words. The word lists
consisted of 120 sentences, and scrambled lists of the same words. Each word was presented on the computer screen for 351ms on average (min: 300ms, max: 1400ms).
Successive words were separated by a blank screen for 300ms, and successive
sentences were separated by an empty screen for a few (3-4) seconds.

\subsection{MEG preprocessing}

The raw MEG data was bandpass-filtered between 0.1 and 40 Hz using MNE-Python default
parameters \cite{mne}. Specifically, we used a zero-phase finite impulse response filter (FIR) with a hamming window and with transition bands of 0.1 Hz and 10 Hz for the low and high
cut-off frequencies respectively.

The raw data was then segmented 100 ms before word onset and 1.000 sec after
word onset. Time $t=0$ms corresponds to the word onset. Finally, each resulting segment was baseline-corrected between -100ms and 0ms, and decimated by 5 and thus led a sampling frequency of 240 Hz. We show, in Figure \ref{fig:megavg}, the average response of the
magnetometers evoked by the words presented to the first subject.

For each word, each subject, and each time sample relative to word onset, we build an observation matrix $Y \in \mathbb{R}^{n, q}$ of $n\approx$ 2700 words by $q=301$ MEG channels (273 magnetometers and 28 compensation channels). Each of the columns of $Y$ is normalized to have a mean and a standard deviation of 0 and 1 respectively.

@@

\subsection{Word features}
We annotated each word with 38 features: 26 corresponds to the number of each letter in the word (from 'a' to 'z', independently of their case), the total number of letters ('word length'), the frequency of the word in natural language (using the Zipf logarithmic scale \cite{marc_brysbaert}), and the part-of-speech tags (i.e. NOUN, PROPN (proper noun), VERB, PRON (pronoun), CONJ (conjunction), DET (determinant), ADJ (adjective), ADV (adverb), and NUM (numerical)) using the packages Spacy \cite{spacy} and WordFreq \cite{speer}. For clarity purposes, we refer to these features using two broad categories: visual (i.e. each letter and the total number of letters) and lexical (part-of-speech and word frequency).
This procedure yields an $X \in \mathbb{R}^{n, d}$ matrix of $n\approx$ 2700 words by $d=38$ features for each subject. Each of the columns of $X$ is normalized to have a mean and a standard deviation of 0 and 1 respectively.

\subsection{Analyses}
We aim to identify which factors of $X$ explain the MEG recordings $Y$ at each time sample and for each subject separately. To this aim, we implemented three types of analyses (forward, backward and back-to-back regressions), and tested whether their results were consistent across subjects, using a second-level statistics described below.

Forward regression, also known as 'encoding' in neuroimaging, here consisted in fitting a Ridge-regularized OLS from $X$ to $Y$:
\begin{equation}
H = (X^{T}X+\lambda I)^{-1} X^{T}Y
\end{equation}
where $\lambda$ is the regularization parameter fit with the leave-one-out cross-validation implemented in scikit-learn \cite{sklearn} from 20 possible value logarithmatically distributed between $10^{-5}$ and $10^5$. We aim to estimate the contribution $diag(\hat E) \in\mathbb{D}^{d} $ of each factor. However, $H\in\mathbb{R}^{d, q}$ is a matrix of $d$ word features by $q$ MEG channels. We thus approximate the relative contribution of each feature by sum-squarring the H coefficients:
\begin{equation}
\hat e_d = \sum_{i=0}^{i=q} {h_d^i}^2
\end{equation}

Backward regression, also known as 'decoding' in neuroimaging, here consisted in fitting a Ridge-regularized OLS from $Y$ to $X$:
\begin{equation}
G = (Y^{T}Y+\lambda I)^{-1} Y^{T}X
\end{equation}
where $\lambda$ is the regularization parameter fit with the leave-one-out cross-validation implemented in scikit-learn \cite{sklearn} from 20 possible value logarithmically distributed between $10^{-5}$ and $10^5$. We aim to estimate the contribution $diag(\hat E) \in\mathbb{D}^{d} $ of each factor. Using 10-fold cross-validation, we thus assessed the ability of this ridge regression to predict independent and identically distributed data:
\begin{equation}
\hat X_{test} = Y_{test}\cdot G_{train}
\end{equation}
and assessing, for each factor of X independently, the Pearson correlation coefficient $r$ between $X_{test}$ and $\hat X_{test}$ for each cross-validation split.

Models were fit for each subject and for each time sample independently

Using this dataset, we study which word features cause a brain response detectable
via MEG.

 $\left(n \times 38\right)$-dimensional input matrix $X$,
where $n$ is the number of words for the subject. Each feature in $X$ is centered and
scaled for zero mean and unit variance.
